Learning action models with minimal observability

### 2018-08-06 doc/aij18/aij18.tex
final conclusions

\section{Conclusions}
\label{sec:conclusions}
We presented a novel approach for learning \strips\ action models from examples using classical planning. The approach is flexible to various amount and kind of input knowledge and accepts partially specified action models. Unlike extensive-data ML approaches, our work explores an alternative research direction that exploits logic reasoning to learn sound models from minimal input knowledge. Remarkably, the action models of the {\em gripper}, {\em npuzzle} or {\em visitall} domains were perfectly learned from only observations of the initial and final states.

To the best of our knowledge, this is the first work on learning and evaluating \strips\ action models from state observations, using exclusively an {\em off-the-shelf} classical planner, and tested over a wide range of different domains.  In many applications, the actual actions executed by the observed agent are not available but, instead, the resulting states can be observed. With this regard, our approach can learn from state observations as it broadens the range of application to external observers and facilitates the representation of imperfect observability, as shown in plan recognition \cite{SohrabiRU16}, as well as learning from unstructured data, like state images \cite{AsaiF18}. When action plans are not available, our approach still produces action models that are compliant with the input information. In this case, since learning is not constrained by actions, operators can be reformulated changing their semantics, in which case the comparison with a reference model turns out to be tricky.

When example plans are available, our approach is also able to compute accurate action models from small sets of learning examples. Our experimental results report that \FAMA achieves {\em precision} and {\em recall} scores that are over 0.7 for ten out of the fourteen IPC domains, using only two plan traces (with ten actions each) and in little computation time (less than a second). Our experimental evaluation also shows that, in the particular setting of minimal input data, \FAMA outperforms \ARMS.

On the other hand, this work also introduced a semantic version of the {\em precision} and {\em recall} metrics. These two metrics, widely used in ML, separately measure the soundness and completeness of the learned models, which facilitate the identification of model flaws. Our brand new semantic version of {\em precision} and {\em recall} do not require a GTM and evaluate \strips\ action models with respect to a given set of observations. In addition, when no reformulation occurs, these new metrics behave similarly to their syntactic counterparts.

Generating {\em informative} examples for learning planning action models is still an open issue. Planning actions include preconditions that are only satisfied by specific sequences of actions which have low probability of being chosen by chance~\cite{fern2004learning}. The success of recent algorithms for exploring planning tasks~\cite{FrancesRLG17} motivates the development of novel techniques that enable to autonomously collect informative learning examples. The combination of such exploration techniques with our learning approach is an appealing research direction that opens up the door to the bootstrapping of planning action models.

%In theory, we could implement a third edit operation for {\em substituting} a fluent from a given operator schema. However, and with the aim of keeping a tractable branching factor of the planning instances that result from our compilations, we only implement {\em deletion} and {\em insertion}

>>>>

\section{Conclusions}
\label{sec:conclusions}

In this paper we have presented \FAMA, an approach for learning \strips\ action models based on the compilation of the learning task into a planning task. The distinctive characteristic of \FAMA over other state-of-the-art approaches is its ability to learn from minimal input knowledge and, more particularly, from minimal observability plan traces with no observed actions and no observed intermediate states. By lightening the input constraints, \FAMA opens up the way for action model learning to operate on real-world problems, as opposite to current approaches where the heavy input restrictions have limited their applicability to synthetic benchmarks. Our approach is thus very well suited for learning action models in data-based applications where the only observable information is a possibly incomplete sequence of partially observed states.

Besides its capacity of working with minimal observability, \FAMA is also able to learn from very small amounts of input knowledge, a clear advantage in domains where obtaining enough training samples is difficult or costly. While the experimental evaluation shows in general an exponential increase of the computation time as the number of training traces augments, \FAMA is able to learn action models more accurate than those of \ARMS with very limited input knowledge. Unlike extensive-data ML approaches, our work explores an alternative research direction that exploits logic reasoning to learn sound models from minimal input knowledge.

This great flexibility of \FAMA to different amounts of minimal input knowledge impacts the complexity of the planning tasks. When the length of the input plan traces is unknown, that is, we ignore the number of steps of the decision-making process of the agent, the planning task to be solved becomes a PSPACE-complete scenario. This is one of the reasons that justifies our compilation-to-planning scheme in contrast to the most extended SAT-solving scheme used in many existing learning systems. Additionally, our planning-based solving scheme allows us to leverage {\em off-the-shelf} classical planners and benefit from the multiple advances in classical planning research.

A key contribution of this work is the proposal of two novel metrics to semantically evaluate the learned action models. These two metrics build upon the well-known syntax-based metrics \emph{precision} and \emph{recall}. Our semantic evaluation mitigates the common issue known as \emph{reformulation} that appears when training sets of minimal observability are used. Due to the lack of observable information, \FAMA can learn semantically valid models that are syntactically different from the reference model. The application of the semantic-based precision and recall allows us to assess the validity of the learned models even in domains where a reference model is not available.
%Remarkably, the action models of the {\em gripper-strips}, {\em n-puzzle} and {\em grid-visit-all} domains were perfectly learned from only observations of the initial and final states.

We highlight that the semantic evaluation of \FAMA is done via the same compilation-to-planning scheme that we use to learn the action models. Given that the input of our learning task definition accepts an initial action model of the agent's behaviour, either complete or partially specified, this solving scheme is exploitable to computing a model that follows the initial model and is compliant with a test set of plan traces. In other words, \FAMA is applicable in model reconciliation tasks by defining a distance metric that measures how close the two models are~\cite{KulkarniCZVZK16}.

More importantly, \FAMA is applicable not only in environments where the domain model is unknown but also in environments where the executable actions are unknown as well. This ability  broadens the range of application to goal reasoning tasks such as plan recognition under imperfect observability \cite{SohrabiRU16}, planning for transparency~\cite{MacNallyLRP18} or counterplanning~\cite{PozancoEFB18}. The application of \FAMA to these tasks offers a plan-based solution where the assumption of a known domain model can be removed. In other words, \FAMA opens up the way towards domain-free goal reasoning.

Finally, we would like to add a note on the open issue of generating {\em informative} plan traces for learning planning action models. Planning actions include preconditions that are only satisfied by specific sequences of actions which have low probability of being chosen by chance~\cite{fern2004learning}. The success of recent algorithms for exploring planning tasks~\cite{FrancesRLG17} motivates the development of novel techniques that enable to autonomously collect informative learning examples. The combination of such exploration techniques with our learning approach is an appealing research direction towards the bootstrapping of planning action models.

%When action plans are not available, our approach still produces action models that are compliant with the input information. In this case, since learning is not constrained by actions, operators can be reformulated changing their semantics, in which case the comparison with a reference model turns out to be tricky.

%When example plans are available, our approach is also able to compute accurate action models from small sets of learning examples. Our experimental results report that \FAMA achieves {\em precision} and {\em recall} scores that are over 0.7 for ten out of the fourteen IPC domains, using only two plan traces (with ten actions each) and in little computation time (less than a second). Our experimental evaluation also shows that, in the particular setting of minimal input data, \FAMA outperforms \ARMS.

%On the other hand, this work also introduced a semantic version of the {\em precision} and {\em recall} metrics. These two metrics, widely used in ML, separately measure the soundness and completeness of the learned models, which facilitate the identification of model flaws. Our brand new semantic version of {\em precision} and {\em recall} do not require a GTM and evaluate \strips\ action models with respect to a given set of observations. In addition, when no reformulation occurs, these new metrics behave similarly to their syntactic counterparts.
